Red Wine Quality Data Analysis by Paul Hsiao
========================================================

This analysis is focused on the red wine dataset.  The red Wine dataset contains 1,599 red wines with 11 variables on the chemical properties of the wine. At least 3 wine experts rated the quality of each wine, providing a rating between 0 (very bad) and 10 (very excellent).

My original intention for this data analysis is to find out whether there is a direct relationship between quality and its chemical contents.  However, the list of chemical included in the dataset missed out a few important ingredients, such as Phenols and Glycerol.  These missed ingredient have big impact on quality of wine, such as smell.  Due to this reason, the focus of this analysis has shifted to finding out the correlation between ingredients and quality in the dataset.

I'll start with univariate analysis to study the distribution of each feature in the dataset. Pick up some features from summary of each feature for further investigation during bivariate analysis and multivariate analysis.  Hopefully, I can come up with a good quality wine profile at the end of the analysis based on the information available.

***

Before jumping into analysis, lets take a look of the dataset first.

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(GGally)
library(scales)
library(memisc)
library(gridExtra)
library(grid)
library(RColorBrewer)
library(bitops)
library(RCurl)
library(ggplot2)
library(dplyr)
```

```{r Load_the_Data, echo=FALSE}
# Load the Data

reddf <- read.csv('wineQualityReds.csv')
reddf$quality <- factor(reddf$quality)
# To avoid inclding the pH.bucket variable used in the scatterplot matrices, a temp variable is used to store the value. The variable is stored in dataset later on.
tmp_pH.bucket <- cut(reddf$pH, breaks = c(2.74, 3.0, 3.25, 3.5, 3.75, 4.01))

```

## Number of observations and variables in the red wine dataset
```{r num_oberv, echo=FALSE}
dim(reddf)
```

## Names of variables for the red wine dataset
```{r list_varname, echo=FALSE}
names(reddf)
```

## Statistics summary of the red wine dataset
```{r show_summary, echo=FALSE}
summary(reddf)
```

##Key observations from the summary
1. fixed.acidity ranges from 4.60 to 15.90 with median 7.90.  Later in the bivariate analysis, I'd like to check what range the high quality wines have.

2. There are observations with zero level citric.acid.  Is this ingredient not necessary to make good wine?  How come it disappeared from some wines?  I'll dig into this issue later on.

3. residual.sugar has max level 15.50 while 3rd Quantile has 2.6 only.  So majority of red wine in the dataset are low in residual.sugar.  I'm curious which quality of wines having high residual.sugar.  I'll find it out later.

4. chlorides has max value 0.611 which is more than 6 times of 3rd quantile value(0.9978). I'd like to see what quality of wines posses these high level of chlorides in bivariate analysis.

5. total.sulfur.dioxide has max value 289.00 while 3rd quantile value is 62.00 only.  I'm also interested to see what quality of wines posses these high level of total.sulfur.dioxide.

***

# Univariate Plots Section

###Histograms for every feature and its summary statistics

```{r Univariate_Plots, echo=FALSE, message=FALSE, warning=FALSE}
# fixed.acidity   
p1 <- ggplot(aes(x=fixed.acidity), data=reddf) +
  geom_histogram(color = 'black', fill = "#FFFF66") +
  scale_x_continuous(breaks = seq(4, 16, 2))

p1

summary(reddf$fixed.acidity)

# volatile.acidity     
p2 <- ggplot(aes(x=volatile.acidity), data=reddf) +
  geom_histogram(color = 'black', fill = "#CCCCCC") +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2))

p2
summary(reddf$volatile.acidity)

# citric.acid
p3 <- ggplot(aes(x=citric.acid ), data=reddf) +
  geom_histogram(color = 'black', fill = "#FF9900") 

p3
summary(reddf$citric.acid)

# residual.sugar       
p4 <- ggplot(aes(x=residual.sugar), data=reddf) +
  geom_histogram(color = 'black', fill = "#CCFF66") +
  scale_x_continuous(limits = c(0, 5))

p4
summary(reddf$residual.sugar)

# chlorides            
p5 <- ggplot(aes(x=chlorides), data=reddf) +
  geom_histogram(color = 'black', fill = "#FFCCCC") 

p5
summary(reddf$chlorides)

# free.sulfur.dioxide  
p6 <- ggplot(aes(x=free.sulfur.dioxide ), data=reddf) +
  geom_histogram(color = 'black', fill = "#0099FF") 

p6
summary(reddf$free.sulfur.dioxide)

# total.sulfur.dioxide
p7 <- ggplot(aes(x=total.sulfur.dioxide), data=reddf) +
  geom_histogram(color = 'black', fill = "#FFCC00")

p7
summary(reddf$total.sulfur.dioxide)

# density              
p8 <- ggplot(aes(x=density), data=reddf) +
  geom_histogram(color = 'black', fill = "#FF66CC") +
  scale_x_continuous(breaks = seq(0.990, 1.004, 0.004))

p8
summary(reddf$density)

# pH                   
p9 <- ggplot(aes(x=pH), data=reddf) +
  geom_histogram(color = 'black', fill = "#00CCFF")

p9
summary(reddf$pH)

# sulphates            
p10 <- ggplot(aes(x=sulphates), data=reddf) +
  geom_histogram(color = 'black', fill = "#666699")

p10
summary(reddf$sulphates)

# alcohol             
p11 <- ggplot(aes(x=alcohol), data=reddf) +
  geom_histogram(color = 'black', fill = "#CCCC33")

p11
summary(reddf$alcohol)

# quality  
p12 <- ggplot(aes(x=quality ), data=reddf) +
  geom_histogram(color = 'black', fill = "#0033CC")

p12

grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12, ncol=4)

```

These histograms show the distribution of each feature in the dataset. The final grid plot provides a summary.  We notices there are a few features showing 'skewed right' histograms, such as total.sulfur.dioxide, chlorides, alcohol, etc.  That might suggest the wine makers want to control the feature contents under certain level.

***

## Re-examine the histograms using different x-scale

Here I use different x-scale to plot the histogram for the same attribute, fixed.acidity. I wonder whether different x-scale can expose some special distribution information.  Let's have a look of the plots.

```{r different_x_scale, echo=FALSE, message=FALSE, warning=FALSE}

# fixed.acidity

p1_1 <- p1 + ggtitle("Histogram for fixed.acidity")
p1_2 <- ggplot(aes(x=fixed.acidity), data=reddf) +
  geom_histogram(color = 'black', fill = "#FF6666") +
  scale_x_log10(breaks = seq(1,20,1)) +
  ggtitle("Histogram for fixed.acidity with scale_x_log10")

p1_3 <- ggplot(aes(x=fixed.acidity), data=reddf) +
  geom_histogram(color = 'black', fill = "#66FF00") +
  scale_x_sqrt(breaks = seq(4, 16, 1)) +
  ggtitle("Histogram for fixed.acidity with scale_x_sqrt")

grid.arrange(p1_1, p1_2, p1_3, ncol=1)

```

These three plots shows different shape of distribution due to different x scale is taken.  The plot with log10 on x-axis gives a more normalized shape.

***

# Univariate Analysis

### What is the structure of your dataset?
```{r echo=FALSE}
str(reddf)
```
All the features have numberic data type.  Here I transformed the quality into factor for plotting purpose.


### What is/are the main feature(s) of interest in your dataset?

fixed.acidity, residual.sugar and density

These three attributes have influence on the acid taste, sweet taste and color which affects the popularity of the wine. 


### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

1. pH - pH should have strong negative correlation with fixed.acidity.
2. alcohol - Alcohol is converted from sugar through fermentation process.


### Did you create any new variables from existing variables in the dataset?

Yes, I created two new variables.: 

1. **pH.bucket**

This variable is created so it can be used as factor which is needed for certain ploting.

2. **total.acidity**

total.acidity = fixed.acidity + volatile.acidity + citric.acid

Summary of total.acidity

```{r total_acidity, echo=FALSE, message=FALSE, warning=FALSE}

reddf$total.acidity <- reddf$fixed.acidity + reddf$volatile.acidity + 
  reddf$citric.acid
summary(reddf$total.acidity)

p13 <- ggplot(aes(x=total.acidity), data=reddf) +
  geom_histogram(color = 'black', fill = "#6600FF") +
  scale_x_continuous(breaks = seq(0, 17, 1))

p13 + ggtitle("Histogram for total.acidity")

```

###Compare the histograms of total.acidity and other acidity features

```{r compare_acidity, echo=FALSE, message=FALSE, warning=FALSE}
grid.arrange(p1, p2, p3, p13, ncol=1)
```

Here we see the fixed.acidity seems to have major influence on the total.acidity which is
not a surprise.  The quantity for volatile.acidity and citric.acid are much smaller compared to fixed.acidity.  So, fixed.acidity has very high correlation to the total.acidity.

How about the relationship between acidity and pH value?  I'll visit this topic in bivariate analysis.

I also want to see the whether the relationship between total acidity and quality can be correlated better than the given acidity features.

Further discussion for these two variables will be included in bivariate analysis section.

***

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?


####Unusual distributions

There are two features that posses unusual distributions:

**1. citric.acid**

citric.acid summary and lowest 6 citric.acid level count:
```{r citric_acid_level_count, echo=FALSE}
summary(reddf$citric.acid)
head(table(reddf$citric.acid))
```

For citric.acid, there are 132 observations having 0 value (out of total 1599 observations).

According to web site Calwineries article (http://www.calwineries.com/learn/wine-chemistry/wine-acids/citric-acid),

"Citric acid plays a major role in the biochemisty of the grape vine and bacteria.... However, its role in wine is far less important.  Most, if not all of the citric acid naturally present in the grapes is consumed by bacteria during fermentation."

"The major disadvantage of adding citric acid is its microbial instability."

This explains why there are 0-level citric.acid.

**2. chlorides**

chlorides summary:

```{r chlorides_check, echo=FALSE}
summary(reddf$chlorides)
```

Majority of observations have chlorides level falls under 0.09.  However, the maximum chlorides level reached 0.611 which is more than 7 times of the median.  This could indicate there are some outliers to cause the high maximum chlorides value.  We'll look at this outlier further later in the bivariate analysis.

***

####Operations on the data to tidy, adjust, or change the form of the data

Due to the nature of quality variable, I transformed the quality variable from numeric to factor.  I did not do any other tidy nor adjust because it is not necessary for my analysis.

***

# Bivariate Plots Section

## Scatterplot matrices

First of all, we'll have a look of the Scatterplot matrices of all features.  This matrices will provide us a rough idea about correlation between all features.  To make the scatterplot matrices more readable, variables are renamed to shorter names for this scatterplot matrices.  The Scatterplot matrices is also stored in a separate file (red_wine_Scatterplot_matrices.JPG) for better viewing.


```{r Bivariate_Plots, echo=FALSE, message=FALSE, warning=FALSE,fig.height=9, fig.width=9}

reddf1 <- reddf
cols <- c("Indx", "f_acid", "v_acid", "c_acid", "sugar", "CL", "f_SO2", "t_SO2", 
          "density", "pH", "S", "alcohol", "quality", "t_acid")
colnames(reddf1) <- cols

ggpairs(reddf1, 
        2:12, 
        axisLabels = "none",
        diag = list(continuous = "bar"),
        title = "Scatter plot matrices for red wine",
        params = c(shape = I('.'), outlier.shape = I('.')))
```

**Abbreviation:**  
   f_acid - fixed.acidity,  
   v_acid - volatile.acidity,  
   c_acid - citric.acid,  
   sugar - residual.sugar,  
   CL - chlorides,  
   f_SO2 - free.sulfur.dioxide,  
   t_SO2 - total.sulfur.dioxide,  
   S - sulphates

From the Scatterplot Matrices, the main observations are:

1. fixed.acidity and density have a positive Correlation Coefficient of 0.668.  This is a interesting point worthy further investigation.  According to web site 'http://winefolly.com/tutorial/how-to-taste-red-wine/', for red wine, higher density means more material is extracted from the grape skin into the wine. That will also lead to darker color.

2. fixed.acidity and pH have a negative Correlation Coefficient -0.683. This is expected as higher acidity will decrease the pH value.

3. fixed.acidity and citric.acid have positive Correlation Coefficient of 0.672. As citric.acid is part of fixed.acidity, the high positive correlation is reasonable.

***

## Revisit histograms with Quality composition revealed

Here, we'll revisit the histogram of a few features.  But this time, we'll show composition of each quality.  With these plots, we can see the influence of level for each feature on the quality.

```{r stack_histogram, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=fixed.acidity, fill = quality), data=reddf) +
  geom_histogram(color = 'black') +
  scale_x_continuous(breaks = seq(4, 16, 1))

ggplot(aes(x=residual.sugar, fill = quality), data=reddf) +
  geom_histogram(color = 'black') +
  scale_x_continuous(limits = c(0, 5))

ggplot(aes(x=chlorides, fill = quality), data=reddf) +
  geom_histogram(color = 'black')

ggplot(aes(x=total.sulfur.dioxide, fill = quality), data=reddf) +
  geom_histogram(color = 'black')

ggplot(aes(x=density, fill = quality), data=reddf) +
  geom_histogram(color = 'black') +
  scale_x_continuous(breaks = seq(0.990, 1.004, 0.002))

ggplot(aes(x=pH, fill = quality), data=reddf) +
  geom_histogram(color = 'black')

ggplot(aes(x=sulphates, fill = quality), data=reddf) +
  geom_histogram(color = 'black')

ggplot(aes(x=alcohol, fill = quality), data=reddf) +
  geom_histogram(color = 'black')
```

As observed, the high quality wines mostly have its composition of each feature in certain range.  The high quality wines rarely have feature contents in highest or lowest range.

***

## Check correlation using scatterplots

Next, lets have a look of correlation between certain features.  The features I'll explore are 
1. fixed.acidity vs pH
2. fixed.acidity vs density
3. alcohol vs residual.sugar

In each plot, subset of 99% quantile of each feature is taken to exclude the 1% outliers. However, full data set is used to calculate the correlation.


### fixed.acidity vs pH

```{r correlation_section1, echo=FALSE, message=FALSE, warning=FALSE}

subdf <- subset(reddf, fixed.acidity <= quantile(reddf$fixed.acidity, 0.99) &
                       pH <= quantile(reddf$pH, 0.99))

ggplot(aes(x = fixed.acidity, y = pH), data = subdf) +
  ggtitle("Jitter scatterplot with smoother, alpha=0.5") +
  geom_jitter(colour="#006633", size = 2, alpha=0.5) +
  geom_smooth(colour="#FF3300")

sprintf("The correlation between fixed.acidity and pH is %s", 
        cor(reddf$fixed.acidity, reddf$pH))
```

This plot clearly shows the fixed.acidity has a negative correlation with pH.

***

### fixed.acidity vs density

```{r correlation_section2, echo=FALSE, message=FALSE, warning=FALSE}

subdf <- subset(reddf, fixed.acidity <= quantile(reddf$fixed.acidity, 0.99) &
                       density <= quantile(reddf$density, 0.99))

ggplot(aes(x = fixed.acidity, y = density), data = subdf) +
  ggtitle("Jitter scatterplot with smoother, alpha=0.5") +
  geom_jitter(color="#66CCFF", size = 2, alpha=0.5) +
  geom_smooth(color="#FF3300")

sprintf("The correlation between fixed.acidity and density is %s", 
        cor(reddf$fixed.acidity, reddf$density))
```

This plot shows a positive correlation between fixed.acidity and density.

***

### alcohol vs residual.sugar

```{r correlation_section3, echo=FALSE, message=FALSE, warning=FALSE}

subdf <- subset(reddf, alcohol <= quantile(reddf$alcohol, 0.99) &
                       residual.sugar <= quantile(reddf$residual.sugar, 0.99))

ggplot(aes(x = alcohol, y = residual.sugar), data = subdf) +
  ggtitle("Jitter scatterplot with smoother, alpha=0.5") +
  geom_jitter(color="#6600FF", size = 2, alpha=0.5) +
  geom_smooth(color="#FF3300")

sprintf("The correlation between alcohol and residual.sugar is %s", 
        cor(reddf$alcohol, reddf$residual.sugar))
```

This plot shows low correlation between alcohol and residual.sugar.  After fermentation process, most of sugar has been transformed into alcohol.  Wine makers have tried to control the residual sugar level to adjust the sweet taste of wine.


These scatterplots with smoother have reinforce the earlier observations from Scatterplot Matrices. The correlation between alcohol vs residual.sugar is low, while it is high for the other pairs of features.

***

## Explore data using boxplots

Next, lets examine the boxplots of a few pair of interested features. We can observe the quantile 1, quantile 3, median and mean value.  After each boxplot, the statistics grouped by quality are listed to provide more precise figures for the boxplot.

```{r boxplot_section, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(aes(x = quality, y =fixed.acidity, fill = quality),
       data = subset(reddf, !is.na(quality))) + geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape =4) +
  ggtitle("Boxplot of fixed.acidity vs quality with mean marked")

reddf.by_quality <- reddf %>%
  group_by(quality) %>%
  summarise(Min = min(fixed.acidity),
            Median = median(as.numeric(fixed.acidity)),
            Mean = mean(fixed.acidity),
            Max = max(fixed.acidity),
            n = n()) %>%
  arrange(quality)

print(reddf.by_quality)

ggplot(aes(x = quality, y =pH, fill = quality),
       data = subset(reddf, !is.na(quality))) + geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape =5) +
  ggtitle("Boxplot of pH vs quality with mean marked")

reddf.by_quality <- reddf %>%
  group_by(quality) %>%
  summarise(Min = min(pH),
            Median = median(as.numeric(pH)),
            Mean = mean(pH),
            Max = max(pH),
            n = n()) %>%
  arrange(quality)

print(reddf.by_quality)

ggplot(aes(x = quality, y =residual.sugar, fill = quality),
       data = subset(reddf, !is.na(quality))) + geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape =4) +
  coord_cartesian(ylim = c(1, 6)) +
  ggtitle("Boxplot of residual.sugar vs quality with mean marked")

reddf.by_quality <- reddf %>%
  group_by(quality) %>%
  summarise(Min = min(residual.sugar),
            Median = median(as.numeric(residual.sugar)),
            Mean = mean(residual.sugar),
            Max = max(residual.sugar),
            n = n()) %>%
  arrange(quality)

print(reddf.by_quality)

ggplot(aes(x = quality, y =alcohol, fill = quality),
       data = subset(reddf, !is.na(quality))) + geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape =3) +
  ggtitle("Boxplot of alcohol vs quality with mean marked")

reddf.by_quality <- reddf %>%
  group_by(quality) %>%
  summarise(Min = min(alcohol),
            Median = median(as.numeric(alcohol)),
            Mean = mean(alcohol),
            Max = max(alcohol),
            n = n()) %>%
  arrange(quality)

print(reddf.by_quality)

ggplot(aes(x = quality, y =chlorides, fill = quality),
       data = subset(reddf, !is.na(quality))) + geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape =4) +
  coord_cartesian(ylim = c(0, 0.2)) +
  ggtitle("Boxplot of chlorides vs quality with mean marked")

reddf.by_quality <- reddf %>%
  group_by(quality) %>%
  summarise(Min = min(chlorides),
            Median = median(as.numeric(chlorides)),
            Mean = mean(chlorides),
            Max = max(chlorides),
            n = n()) %>%
  arrange(quality)

print(reddf.by_quality)

ggplot(aes(x = quality, y =total.sulfur.dioxide, fill = quality),
       data = subset(reddf, !is.na(quality))) + geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape =2) +
  coord_cartesian(ylim = c(0, 200)) +
  ggtitle("Boxplot of total.sulfur.dioxide vs quality with mean marked")

reddf.by_quality <- reddf %>%
  group_by(quality) %>%
  summarise(Min = min(total.sulfur.dioxide),
            Median = median(as.numeric(total.sulfur.dioxide)),
            Mean = mean(total.sulfur.dioxide),
            Max = max(total.sulfur.dioxide),
            n = n()) %>%
  arrange(quality)

print(reddf.by_quality)

ggplot(aes(x = quality, y =density, fill = quality),
       data = subset(reddf, !is.na(quality))) + geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape =3) +
  ggtitle("Boxplot of density vs quality with mean marked")

reddf.by_quality <- reddf %>%
  group_by(quality) %>%
  summarise(Min = min(density),
            Median = median(as.numeric(density)),
            Mean = mean(density),
            Max = max(density),
            n = n()) %>%
  arrange(quality)

print(reddf.by_quality)

```

These boxplots show some interesting information for the wine with high rating quality.  They demonstrates the distribution of the focused features per quality.  The following are observed for the quality 8 wines:

   a. Low pH value              (median: 3.23)
   b. Low residual sugar        (median: 2.1)
   c. High alcohol              (median: 12.15)
   d. Low chlorides             (median: 0.0705)
   e. Low total.sulfur.dioxide  (median: 21.5)
   f. Low density               (median: 0.994940)

***

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

From the Scatterplot matrices, the following are observed:

1. fixed.acidity and density have positive Correlation Coefficient of 0.668.

2. fixed.acidity and pH have negative Correlation Coefficient -0.683. This is expected as higher acidity will decrease the pH value.

3. The rest of the correlations are weak. None of the features has high correlation with quality directly.

4. The scatterplots with smoother have reinforced the above observations.

***

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

free.sulfur.dioxide and total.sulfur.dioxide have a positive Correlation Coefficient of 0.668. However, this is not a surprise as free.sulfur.dioxide is included in total.sulfur.dioxide.

***

### What was the strongest relationship you found?

fixed.acidity and pH have the highest Correlation Coefficient -0.683.  Wine makers can adjust the pH level, ie. fixed acidity level, so their wine can satisfy consumer's taste.  In this dataset, the median pH level is 3.310.

***

# Multivariate Plots Section

## 1. pH vs median alcohol per quality

First, lets have a look of pH vs median alcohol per quality.

```{r Multivariate_Plot1, echo=FALSE, message=FALSE, warning=FALSE}

reddf$pH.bucket <- tmp_pH.bucket

alcohol_by_pH_quality <- reddf %>%
  filter(!is.na(quality)) %>%
  group_by(pH, quality) %>%
  summarise(median_alcohol = median(as.numeric(alcohol)),
            n = n()) %>%
  ungroup() %>%
  arrange(pH)

ggplot(aes(x = pH, y =median_alcohol),
       data = alcohol_by_pH_quality) + 
  geom_point(aes(color = quality)) +
  geom_smooth(aes(color = quality), size=1.5) +
  ggtitle("Plot 3-1: pH vs median_alcohol with smoother")

sprintf("The correlation between median_alcohol and pH is %s", 
        cor(alcohol_by_pH_quality$median_alcohol, alcohol_by_pH_quality$pH))

```

This plot shows the pH has a positive correlation with median alcohol value for each quality.  For higher quality wines, the median alcohol tends to have higher value for the same pH value.  This plot gives better details into each quality than the following plot which shows the smooth trend across all qualities of wine.


```{r Multivariate_Plot2, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(aes(x = pH, y = median_alcohol),
       data = alcohol_by_pH_quality) + 
  geom_point(aes(color = quality)) +
  geom_smooth(size=1.5) +
  ggtitle("Plot 3-2: pH vs median_alcohol with smoother across all qualities")
```

***

## 2. alcohol vs density per pH bucket

Next, lets look at the alcohol vs density per pH bucket.

```{r Multivariate_Plot3, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(aes(x = alcohol, y = density),
       data = subset(reddf, !is.na(pH.bucket))) + 
  geom_point(aes(color = pH.bucket)) +
  geom_smooth(aes(color = pH.bucket), size=1.5) +
  ggtitle("Plot 3-3: alcohol vs density per pH.bucket") 

sprintf("The correlation between alcohol and density is %s", 
        cor(reddf$density, reddf$alcohol))

```

This plot shows a confusing trends, especially for the pH.bucket 2.7-4.3.  Could this plot be presented better if we remove some outliers?  Let's have a look of the distribution of density per pH.bucket:

```{r Multivariate_data1, echo=FALSE, message=FALSE, warning=FALSE}
reddf.by_ph_bucket1 <- reddf %>%
  filter(!is.na(pH.bucket)) %>%
  group_by(pH.bucket) %>%
  summarise(Min = min(density),
            Median = median(as.numeric(density)),
            Mean = mean(density),
            Max = max(density),
            n = n()) %>%
  arrange(pH.bucket)

print(reddf.by_ph_bucket1)
```

And here is the distribution of alcohol per pH.bucket:

```{r Multivariate_data2, echo=FALSE, message=FALSE, warning=FALSE}
reddf.by_ph_bucket <- reddf %>%
  filter(!is.na(pH.bucket)) %>%
  group_by(pH.bucket) %>%
  summarise(Min = min(alcohol),
            Median = median(as.numeric(alcohol)),
            Mean = mean(alcohol),
            Max = max(alcohol),
            n = n()) %>%
  arrange(pH.bucket)

print(reddf.by_ph_bucket)
```


Now, I'll show the plot with top 1% alcohol and density removed from the dataset, plus jitter and alpha=0.5.

```{r Multivariate_Plot4, echo=FALSE, message=FALSE, warning=FALSE}

reddf1 <- subset(reddf, 
                 alcohol <= quantile(reddf$alcohol, 0.99) &
                 alcohol >= quantile(reddf$alcohol, 0.01) &
                 density <= quantile(reddf$density, 0.99) &
                 density >= quantile(reddf$density, 0.01))

ggplot(aes(x = alcohol, y = density),
       data = subset(reddf1, !is.na(pH.bucket))) + 
  geom_jitter(aes(color = pH.bucket), alpha=0.5) +
  geom_smooth(aes(color = pH.bucket), size=1.5) +
  ggtitle("Plot 3-4: alcohol vs density per pH.bucket with 1% outlier excluded")

sprintf("The correlation between alcohol and density is %s", 
        cor(reddf1$density, reddf1$alcohol))

```

This plot looks much better compared with previous plot.  The trend for all pH.buckets are obvious, except the pH.bucket 3.75-4.01.  This is due to pH.bucket 3.75-4.01 has small number of sample data.  In general, the alcohol has a negative correlation with density.

## 3. Build a model

Now, lets try to build a model to show some statistical measures.

```{r Multivariate_model, echo=FALSE, message=FALSE, warning=FALSE}

# to be able to use quality in the model, need to transfer from factor back to numeric
reddf$quality <- as.numeric(levels(reddf$quality))[reddf$quality]

m1 <- lm(I(quality) ~ I(pH), data = reddf)
m2 <- update(m1, ~ . + density)
m3 <- update(m2, ~ . + fixed.acidity)
m4 <- update(m3, ~ . + alcohol)
m5 <- update(m4, ~ . + residual.sugar)
mtable(m1, m2, m3, m4, m5)

# need to transform the quality to factor for boxplot
reddf$quality <- factor(reddf$quality)

```

The result from this linear model seems indicate the fixed.acidity and alcohol have more impact to the quality figure than other feature in the model.

***

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?
From plot 3-1 and 3-2, it is observed the pH and median_alcohol has positive correlation. Higher quality wines have higher median_alcohol value.  So, the pH and alcohol are strengthened each other in this analysis.

From plot 3-4, it shows the alcohol and density has negative correlation.

***

### Were there any interesting or surprising interactions between features?

In plot 3-3 and 3-4, the line for lowest pH.bucket are in big density range and accross the full alcohol range. For this lowest pH.bucket line, when it comes to high alcohol range, the density comes down to it lowest range. Compare this observation with previous observations from boxplot, they matches with the observatons from quality 8.

***

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

Yes, I did create a model. The strength is it has low p value. However, the R-squared is quite low.

------

# Final Plots and Summary

## Plot One
```{r Plot_One, echo=FALSE, message=FALSE, warning=FALSE }
grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12, ncol=4)
```

## Description One

This is the grid of histograms of all features.  It show the distribution of each feature in the dataset. It provides the general idea about the dataset variable distribution before we looking into correlation between features.

***

## Plot Two
```{r Plot_Two, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = quality, y =alcohol, fill = quality),
       data = subset(reddf, !is.na(quality))) + geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape =3) +
  ggtitle("Boxplot of alcohol vs quality with mean marked")
```

## Description Two

This plot clearly demonstrates the relationship between alcohol and quality.  Even though the quality depends on other features as well, the plot shows the higher alcohol level is correlated to higher quality value in general.

***

## Plot Three
```{r Plot_Three, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = pH, y =median_alcohol),
       data = alcohol_by_pH_quality) + 
  geom_point(aes(color = quality)) +
  geom_smooth(aes(color = quality), size=1.5) +
  ggtitle("pH vs median_alcohol with smoother")
```

## Description Three

This plot shows the pH vs median_alcohol trend -  
1. For the same pH level, higher median_alcohol level linked to higher quality value.  
2. For the same median_alcohol level, lower pH level leads to higher quality value. 

------

# Reflection

It is a interesting subject to explore using the Data Analysis we learnt in this course. Without knowing much about the relationship between features, it was a bit hard to pick the features of interest in the begining.  However, once I went through the Univariate, bivariate and multivariate ploting involving different feature combinations, the picture is getting clear. Even though there is no definite conclusion from the analysis, a few key observations did provide us a good picture for the relationship of quality and interested features.  There are still a lot of plotting technics we have not applied during the course yet.  Even though the dataset provides only limited features in red wine, I believe with further exploring using more plotting technics, we can be more certain about the relationship between quality and features.
